\chapter*{Acknowledgements}


First and foremost, I would like to thank God. Without his blessings, none of these achievements would have been possible.

This work would not have been possible without the support of my supervisor, Tobias Ludwig. Thank you for your help, time, and valuable advice. I would like to express my gratitude and appreciation to coordinator of the EMECS program, Prof. Wolfgang Kunz. Your dedication to run this program does not go unnoticed. I would like to extend my acknowledgement to all the EMECS professors and administrators from the University of Kaiserslautern (TUK) and Norwegian University of Science and Technology (NTNU) . A special thank you to all the staff of the Student Office from NTNU and ISGS from TUK, you all made the process much easier. 

I would also like to express my gratitude to my family and all my friends. I am very fortunate to have so many of you that I cannot personally mention everyone. This path would have not been the same without each of you.

I have also to acknowledge all professors that I had the pleasure to meet, study and work with so far. I admire your work, dedication, and contribution not only to academy, but specially to society. You are an inspiration to me, and you are making a better world. 

A huge and special thank to my brother, Sidney. You walked this path before me and showed me that it is possible. You are my role model. 

Finally, I would like to express my enormous gratitude and appreciation to my parents, Sérgio and Nora. I could not possibly come this far without your support and love. Thank you for always believing in me even when I could not. You are my heroes.
